---
title: "Clustering Project"
author: "Taweepong Hirikul"
date: '2022-06-09'
output:
  word_document: default
  html_document: default
---

#STEP 1: load & process the data
##1.1 Laod the data
```{r}
df <- read.delim("BC_Educ_Labour_Market_Performance.csv", header = TRUE, sep = ",")
head(df)
```
##1.2 exclude text column
```{r}
df_sel <- df[,-c(1)]
head(df_sel)
```
##1.3 download & library package
```{r}
install.packages("caret")
```

```{r}
library(caret)
```

#STEP 2: Normalize/scale data using min-max normalization.
```{r}
#preproc <- preProcess(df,method = c("center", "scale"), rangeBounds = c(2:24))
#preproc
process <- preProcess(df_sel, method=c("range"))
process
df_scale <- predict(process, df_sel)
df_scale
```

#STEP 3: use normalized data to perform cluster analysis using k-means method.
##3.1 download package
```{r}
install.packages("factoextra")
install.packages("NbClust")
```
##3.2 library package
```{r}
library(factoextra)
library(NbClust)
library(cluster)
```
##3.3 use NbClust
```{r}
res_nbclust <- NbClust(df_scale, distance = "euclidean", min.nc = 2, max.nc = 10, method = "kmeans", index = "all")
fviz(res_nbclust)
```
#STEP 4: determining the appropriate number of clusters to use in the analysis.
Determine for the number of clusters is 3
```{r}
fviz_nbclust(df_scale,kmeans, method="wss")
```


#STEP 5: Perform k-means clustering using the number of clusters you obtained from step 3 above.
```{r}
set.seed(45)
km_res <- kmeans(df_scale, 3, nstart = 50)
print(km_res)
str(km_res) #examine the structure of the result object

```

#STEP 6: Validate clustering results using appropriate techniques.
```{r}

clusplot(df_scale,km_res$cluster, main="2D Cluster Solution Representation",color=TRUE,shade=TRUE, lines=0, labels=0)
```
###silhouette coeffcient (use silhouette function)
```{r}
sil_out <- silhouette(km_res$cluster,dist(df_scale,method="euclidean"))
```

###plot silhouette results (use plot() function)
```{r}
plot(sil_out,main="Silhouette Plot from Kmeans Analysis", col=c("red","green"))
```

#STEP 7: Obtain a summary table of the number of observations assigned to each cluster

###generate a table of cluster membership
###use table() function
```{r}
table(km_res$cluster)
```
###display cluster centers/means as a df
```{r}
df_centers <- data.frame(km_res$centers)
df_centers
```
### transpose cluster centers table
```{r}
df_centers1 <- t(df_centers)
df_centers1
```

##Culster centers using the original data
```{r}
clust_cent_orig <- aggregate(df_sel, list(cluster=km_res$cluster), mean)
clust_cent_orig1 <- t(clust_cent_orig)
clust_cent_orig1

```

#STEP 8: Repeat the analysis using hierarchical clustering technique and generate the dendogram.

```{r}
dmax <- dist(df_sel, method = "euclidean")
h_hclust <- hclust(dmax, method = "ward.D")
plot(h_hclust, cex=0.5,hang = -1, main = "Dendogram from Hclust Algorithm")
```

```{r}
fviz_dend(h_hclust,k=3,color_labels_by_k = TRUE, cex = 0.5)
```

#STEP 9: Does the rule-of-thumb of determining the number of clusters using the dendogram confirm the number of clusters obtained from k-means method?

YES, The dendogram is confirm that the number of cluster from k-means method is 3!!

#STEP 10: Using the number of clusters determined from k-means method, cut the dendogram/tree. As you did in step 7 above.
```{r}
dmax <- dist(df_sel, method = "euclidean")
h_hclust <- hclust(dmax, method = "ward.D")
#plot(h_hclust, cex=0.5,hang = -1, main = "Dendogram from Hclust Algorithm")
fviz_dend(h_hclust,k=3,color_labels_by_k = TRUE, cex = 0.5)
```
##cut the dendogram (use cutree function)

```{r}
df_cut <- cutree(h_hclust,k=3)
table(df_cut)
```

##generate cluster centroids (means) table (use aggregate function)

```{r}
df_center <- aggregate(df_sel,list(cluster=df_cut),mean)
df_center
```
##transpose the clust_centers (use t function)

```{r}
df_center1 <- t(df_center)
df_center1
```

#STEP 11: Append cluster assignment to the original data file you imported for analysis and save the output in each as a CSV file.
##Append cluster membershilp to the original dataset
```{r}
df$clust_mem <- df_cut
df
```
##Save the output to CSV
```{r}
write.csv(df,"df_with_cluster_mem.csv")
```


